Simulating an Isolated Dwarf Galaxy with Energetic Feedback and Chemical Yields from Individual Stars

### 2018-03-14 main.tex
changes through 2.5.1

%mm2 I think there does need to be some introduction to this discussion:
However, much of this work has been done with Lagrangian smoothed particle hydrodynamics (SPH) schemes, with a few recent exceptions  \citep{Few2012,Few2014,Vorobyov2015}. In its original form, SPH does not capture mixing between chemically inhomogeneous particles, as necessary for chemical evolution. This can be modeled with sub-grid models of turbulent metal diffusion \citep[e.g.][]{Shen2010,Shen2013,Brook2014,Su2017a,Escala2018}, but there are many possibly models that are not necessarily applicable in every regime \citep[see ][]{Revaz2016}. While mixing occurs in Eulerian codes even without sub-grid models, numerical diffusion tends to result in over-mixing, requiring very high resolution or additional viscosity to reduce. Molecular diffusion or even turbulent mixing is certainly not resolved in any galaxy-scale simulation with either method, requiring additional sub-grid models; this can be particularly important for understanding the initial pollution of otherwise pristine gas \citep[see ][ and references therein]{PanScannapiecoScalo2013,Sarmento2017}. Moreover, metal mixing efficiencies may vary species-by-species, as suggested by recent work \citep[e.g.][]{Cohen2013, Roederer2014, FrebelNorris2015, Hirai2017, Cote2017}. Mixing behavior is tied critically to the feedback source (stellar winds, supernovae, and possibly more exotic sources) that inject metals into different phases of 
the ISM with different energies and on different timescales; the observational effect of this is poorly understood, however. The variations in how different methods handle sub-grid metal injection and metal mixing schemes can lead to uncertainties in connecting models to observations and the fundamental physics that drives galaxy evolution.

@@

We refine the mesh whenever the Jeans length is no longer resolved by a minimum of 8 cells, continually refining a given region until this criteria is met or until the region reaches the maximum resolution (1.8~pc). At maximum resolution, the Jeans length can become under resolved, leading to artificial numerical fragmentation. \citet{Truelove1997} showed that resolving the Jeans length by at least 4 cells is required to suppress this fragmentation.

@@

We use the chemistry and cooling library \texttt{Grackle}\footnote{https://grackle.readthedocs.io/en/grackle-3.0/} v 3.0 to follow a nine species non-equilibrium chemistry network (H, H$^+$, He, He$^+$, He$^{++}$, e$^{-}$, H$_2$, H$^{-}$, and H$_{2}$) which includes radiative heating and cooling from these species and metals.\footnote{We use a slightly modified version of the the main \texttt{Grackle} repository, available at https://bitbucket.org/aemerick/grackle-emerick at changeset 59adc76b3130.} \texttt{Grackle} is a freely available, open source, multi-code library, designed to interface with a wide variety of astrophysical codes. We outline specific model choices made in our simulations and refer the reader to \citet{GrackleMethod} for a detailed discussion of the code. We apply the \citet{Glover2008} three-body rate for H$_{2}$ formation and include a model for H$_2$ formation on dust, dust heating, and dust cooling following the methods in \citet{2000ApJ...534..809O} and \citet{2005ApJ...626..627O} as included in \texttt{Grackle}. However, we update the default dust to gas ratio scaling in \texttt{Grackle} to account for the steeper scaling in low metallicity regimes ($Z \lesssim 0.1 Z_{\odot}$) with the broken power law scalings from \citet{Remy-Ruyer2014}. For metallicities above $\sim 0.1 Z_{\odot}$, this is equivalent to the the default behavior of \texttt{Grackle}, where dust content scales linearly with metallicity.
% Since we assume a dust abundance that depends only on metallicity, we do not track dust destruction or creation, which may play an important role regulating detailed ISM properties. (Vogelsberger is working on this directly but no papers yet... will cite if it gets submitted by then)

As part of the \texttt{Grackle} package, metal line cooling is modeled using pre-computed \texttt{Cloudy} \citep{Cloudy2013} \footnote{http://www.nublado.org/} tables interpolated as a function of density, temperature, and redshift, using the \citet{HM2012} UV metagalactic background. As discussed in more detail in Section~\ref{sec:diffusive heating}, we account for approximate self-shielding of H and He against this UV background. Using this prescription with metal line cooling tables computed under an optically thin assumption can lead to an order of magnitude overestimation of the cooling rate at certain densities, as discussed in \citet{Hu2017} and  Appendix~\ref{appendix:cooling}. To address this issue, we use re-computed metal line tables consistent with the self-shielding approximation.  We have made these new tables public in the main \texttt{Grackle} repository. These are discussed in greater detail in Appendix ~\ref{appendix:cooling}. {\bf {\it (addressing MM's comment below (I was unable to find a ref. where this was accounted for)}: Finally, we ignore the effect the stellar radiation field (see Section~\ref{sec:diffusive heating}) may have on the interpolated metal line cooling rates. }

%mm2 
   {\bf [MM: I'm a bit confused here: how does the ISRF interact with the cooling?  Shouldn't that move you around in the Cloudy tables, possibly quite substantially?]}

\subsection{Star Formation Algorithm}
\label{sec:star formation}
In order to resolve individual star formation events on galaxy scales, we implement a stochastic star formation algorithm adopted from \citet{Goldbaum2015,Goldbaum2016}. Each cell at the maximum refinement level is capable of forming stars if it contains gas that meets the following local criteria on number density $n$, temperature $T$, cell mass $M$, and velocity $\vec{v}$: 1) $n > n_{\rm thresh}$, 2) $T < T_{\rm thresh}$, 3) $M > M_{\rm Jeans}$, and 4) $\vec{\nabla} \cdot \vec{v} < 0$, where $n_{\rm thresh}$ is a resolution dependent density threshold, $T_{\rm thresh}$ is a temperature threshold, and $M_{\rm Jeans}$ is the local Jeans mass. Our fiducial values are $n_{\rm thresh} =  200 \mbox{ cm}^{-3}$ and $T_{\rm thresh}= 200$~K. We limit the fraction of a cell's gas mass that is converted into stars by requiring $M > f_{\rm thresh} M_{\rm max,*}$, where $f_{\rm thresh} = 2.0 $ and the maximum star mass $M_{\rm max,*} = 100 \mbox{ M}_{\odot}$. No star formation occurs when $M < f_{\rm thresh} M_{\rm max,*}$, ensuring that a star formation episode can not produce negative densities.

We make the common ansatz that star formation occurs by converting some fraction of gas into stars in a free fall time ($\tau_{\rm ff}$) as characterized by a star formation efficiency, $\epsilon_{\rm f} \simeq 0.02$. At high resolution, one's choice of $\epsilon_{\rm f}$ should be irrelevant \citep{Orr2017, FIRE2}, as star formation is ultimately self-regulated by feedback.
%but we find that $\epsilon_{\rm f}$ may impact star formation in the lowest mass halos, which are unusually sensitive to stellar ionization feedback (see Emerick et. al. in prep).

@@

\subsubsection{Stellar Yields}
\label{sec:yields}
For the first time in galaxy scale simulations, we track galactic chemodynamical evolution using stellar yields as ejected from star particles that represent individual stars. We adopt the NuGrid\footnote{http://www.nugridstars.org} collaboration's set of stellar yields given on a uniformly sampled grid in stellar mass and metallicity with 12 mass bins over $M_{*} \in \left[1.0, 25.0\right]$ M$_{\odot}$ and five metallicity bins at $Z =$ 0.02, 0.01, 0.006, 0.001, and 10$^{-4}$ \citep{Pignatari2016, Ritter2017}. This grid includes yields from the AGB phase of stars from 1--7 M$_{\odot}$, as well as yields from both stellar winds and core collapse SNe of massive stars from 12--25 M$_{\odot}$. We complement these tables with tables from Slemer et. al. (in prep), based on the PARSEC stellar evolution tracks \citep{Bressan2012, Tang2014} to track stellar winds for stars more massive than 25 M$_{\odot}$. We ignore SN yields from these stars (see next paragraph). We combine all stable isotope yields for a given element into a single elemental abundance for all stable elements from hydrogen to bismuth. Although we can in principle follow an arbitrary number of metal species, practical considerations of memory use prevent this in any given simulation. We refer the reader to previous uses of the NuGrid yields in one-zone galactic chemical enrichment models \citep{Cote2016,  Cote2016_feb,Cote2017} for a detailed discussion of how various model uncertainties can influence galactic chemical evolution.

Above some mass $M_{\rm trans}$ within the unsampled range of 7--12 M$_{\odot}$, stars do not undergo AGB wind phases and end their lives as core collapse SNe. Where this transition occurs is uncertain, but is commonly taken to be around 8--10~M$_{\odot}$. This is a free parameter in our model, but we take $M_{\rm trans} = 8$~M$_{\odot}$ as our fiducial value. In our model, stars below this mass eject their wind yields in an AGB phase only at the end of their lives, typically over a period comparable to or less than a few time steps ($\lesssim 10$ kyr). Stars above this mass are assumed to eject their stellar yields via line-driven stellar winds at a constant mass loss rate throughout their lifetime (neglecting Wolf-Rayet and luminous blue variable phases), ending their lives as a core-collapse SN (see Sect.~\ref{sec:stellar winds} for details on the wind energetics). Varying $M_{\rm thresh}$ changes both the time at which yields for stars around this mass are ejected (for reference, the lifetime of an 8 M$_{\odot}$ star is about 35--40~Myr), and the energy injection from these winds. \citet{Cote2017} explores how the choice of $M_{\rm thresh}$ affects galaxy abundances in a one-zone model. We neglect the effects of binary star evolution on stellar feedback, and discuss the significance of this in Sect.~\ref{sec:binary stars}.

@@

Type Ia SNe are an important additional source of galactic chemical enrichment. These iron group rich events are responsible for the $\sim$1 Gyr timescale turnover, or ``knee", in $[\alpha/\rm{Fe}]$ vs $[\rm{Fe}/\rm{H}]$ diagrams. We use the Type Ia SN yields given in \citet{Thielemann1986}, adopting a Type Ia SN model as discussed in Sect.~\ref{sec:Type Ia}. We emphasize that we only track Type Ia SNe occurring within the population of stars formed in this model, neglecting SNe from any possible pre-existing population.

>>>>

%mm2 I think there does need to be some introduction to this discussion:
However, much of this work has been done with Lagrangian smoothed particle hydrodynamics (SPH) schemes, with a few recent exceptions  \citep{Few2012,Few2014,Vorobyov2015}. In its original form, SPH does not capture mixing between chemically inhomogeneous particles, as necessary for chemical evolution. This can be modeled with sub-grid models of turbulent metal diffusion \citep[e.g.][]{Shen2010,Shen2013,Brook2014,Su2017a,Escala2018}, but there are many possibly models that are not necessarily applicable in every regime \citep[see ][]{Revaz2016}. While mixing occurs in Eulerian codes even without sub-grid models, numerical diffusion tends to result in over-mixing, requiring very high resolution
%mm3 or additional viscosity [I'm not sure I understand this.  Adding explicit viscosity will increase mixing, although it at least places it on a physical basis, rather than relying on error terms with unphysical scale dependences.]
to reduce. Molecular diffusion or even turbulent mixing is certainly not resolved in any galaxy-scale simulation with either method, requiring additional sub-grid models; this can be particularly important for understanding the initial pollution of otherwise pristine gas \citep[see ][ and references therein]{PanScannapiecoScalo2013,Sarmento2017}. Moreover, metal mixing efficiencies may vary species-by-species, as suggested by recent work \citep[e.g.][]{Cohen2013, Roederer2014, FrebelNorris2015, Hirai2017, Cote2017}. Mixing behavior is tied critically to the feedback source (stellar winds, supernovae, and possibly more exotic sources) that inject metals into different phases of 
the ISM with different energies and on different timescales; the observational effect of this remains poorly understood, however. The variations in how different methods handle sub-grid metal injection and metal mixing schemes can lead to uncertainties in connecting models to observations and the fundamental physics that drives galaxy evolution.

@@

We refine the mesh whenever the Jeans length is no longer resolved by a minimum of 8 cells, continually refining a given region until this criterion is met or until the region reaches the maximum resolution (1.8~pc). At maximum resolution, the Jeans length can become under-resolved, leading to artificial numerical fragmentation. \citet{Truelove1997} showed that resolving the Jeans length by at least 4 cells is required to suppress this fragmentation.

@@

We use the chemistry and cooling library \texttt{Grackle}\footnote{https://grackle.readthedocs.io/en/grackle-3.0/} v 3.0 to follow a nine species non-equilibrium chemistry network (H, H$^+$, He, He$^+$, He$^{++}$, e$^{-}$, H$_2$, H$^{-}$, and H$_{2}$) which includes radiative heating and cooling from these species and metals.\footnote{We use a slightly modified version of the the main \texttt{Grackle} repository, available at https://bitbucket.org/aemerick/grackle-emerick at changeset 59adc76b3130.} \texttt{Grackle} is a freely available, open source, multi-code library, designed to interface with a wide variety of astrophysical codes. We outline specific model choices made in our simulations and refer the reader to \citet{GrackleMethod} for a detailed discussion of the code. We apply the \citet{Glover2008} three-body rate for H$_{2}$ formation and include a model for H$_2$ formation on dust, dust heating, and dust cooling following the methods in \citet{2000ApJ...534..809O} and \citet{2005ApJ...626..627O} as included in \texttt{Grackle}. However, we update the default dust to gas ratio scaling in \texttt{Grackle} to account for the steeper scaling in low metallicity regimes ($Z \lesssim 0.1 Z_{\odot}$), using the broken power law scalings from \citet{Remy-Ruyer2014}. For metallicities above $\sim 0.1 Z_{\odot}$, this is equivalent to the default behavior of \texttt{Grackle}, where dust content scales linearly with metallicity.
% Since we assume a dust abundance that depends only on metallicity, we do not track dust destruction or creation, which may play an important role regulating detailed ISM properties. (Vogelsberger is working on this directly but no papers yet... will cite if it gets submitted by then)

As part of the \texttt{Grackle} package, metal line cooling is modeled using pre-computed \texttt{Cloudy} \citep{Cloudy2013} \footnote{http://www.nublado.org/} tables interpolated as a function of density, temperature, and redshift, using the \citet{HM2012} UV metagalactic background. As discussed in more detail in Section~\ref{sec:diffusive heating}, we account for approximate self-shielding of H and He against this UV background. Using this prescription with metal line cooling tables computed under an optically thin assumption can lead to an order of magnitude overestimation of the cooling rate at certain densities, as discussed in \citet{Hu2017} and  Appendix~\ref{appendix:cooling}. To address this issue, we use re-computed metal line tables consistent with the self-shielding approximation.  We have made these new tables public in the main \texttt{Grackle} repository. These are discussed in greater detail in Appendix ~\ref{appendix:cooling}.
%mm3 {\bf {\it (addressing MM's comment below (I was unable to find a ref. where this was accounted for)}: 
Finally, we ignore the effect the stellar radiation field (see Section~\ref{sec:diffusive heating}) may have on the interpolated metal line cooling rates. 
%}
%%mm2 
%   {\bf [MM: I'm a bit confused here: how does the ISRF interact with the cooling?  Shouldn't that move you around in the Cloudy tables, possibly quite substantially?]}

\subsection{Star Formation Algorithm}
\label{sec:star formation}
In order to resolve individual star formation events on 
%mm3 galaxy 
    galacitc
scales, we implement a stochastic star formation algorithm adopted from \citet{Goldbaum2015,Goldbaum2016}. Each cell at the maximum refinement level is capable of forming stars if it contains gas that meets the following local criteria on number density $n$, temperature $T$, cell mass $M$, and velocity $\vec{v}$: 1) $n > n_{\rm thresh}$, 2) $T < T_{\rm thresh}$, 3) $M > M_{\rm Jeans}$, and 4) $\vec{\nabla} \cdot \vec{v} < 0$, where $n_{\rm thresh}$ is a resolution dependent density threshold, $T_{\rm thresh}$ is a temperature threshold, and $M_{\rm Jeans}$ is the local Jeans mass. Our fiducial values are $n_{\rm thresh} =  200 \mbox{ cm}^{-3}$ and $T_{\rm thresh}= 200$~K. We limit the fraction of a cell's gas mass that is converted into stars by requiring $M > f_{\rm thresh} M_{\rm max,*}$, where $f_{\rm thresh} = 2.0 $ and the maximum star mass $M_{\rm max,*} = 100 \mbox{ M}_{\odot}$. No star formation occurs when $M < f_{\rm thresh} M_{\rm max,*}$, ensuring that a star formation episode can not produce negative densities.

We make the common ansatz that star formation occurs by converting 
%mm3 some fraction of 
gas into stars in a free fall time $\tau_{\rm ff}$ 
%mm3 characterized by 
   with 
a star formation efficiency, $\epsilon_{\rm f} \simeq 0.02$. At high resolution, 
the choice of $\epsilon_{\rm f}$ should be irrelevant \citep{Orr2017, FIRE2}, as star formation is ultimately self-regulated by feedback.
%but we find that $\epsilon_{\rm f}$ may impact star formation in the lowest mass halos, which are unusually sensitive to stellar ionization feedback (see Emerick et. al. in prep).

@@

\subsubsection{Stellar Yields}
\label{sec:yields}
For the first time in galaxy-scale simulations, we track galactic chemodynamical evolution using stellar yields ejected from star particles that represent individual stars. We adopt the NuGrid\footnote{http://www.nugridstars.org} collaboration's set of stellar yields given on a uniformly sampled grid in stellar mass and metallicity with 12 mass bins over $M_{*} \in \left[1.0, 25.0\right]$ M$_{\odot}$ and five metallicity bins at $Z
%mm3 [confirm]
       / Z_{\odot} 
=$ 0.02, 0.01, 0.006, 0.001, and 10$^{-4}$ \citep{Pignatari2016, Ritter2017}. This grid includes yields from the AGB phase of stars from 1--7 M$_{\odot}$, as well as yields from both stellar winds and core collapse SNe of massive stars from 12--25 M$_{\odot}$. We complement these tables with tables from Slemer et al.\ (in prep), based on the PARSEC stellar evolution tracks \citep{Bressan2012, Tang2014}, to track stellar winds for stars more massive than 25 M$_{\odot}$. We ignore SN yields from these stars (see next paragraph). We combine all stable isotope yields for a given element into a single elemental abundance for all stable elements from hydrogen to bismuth. Although we can in principle follow an arbitrary number of metal species, practical considerations of memory use prevent this in any given simulation. We refer the reader to previous uses of the NuGrid yields in one-zone galactic chemical enrichment models \citep{Cote2016,  Cote2016_feb,Cote2017} for a detailed discussion of how various model uncertainties can influence galactic chemical evolution.

Above some mass $M_{\rm trans}$ within the unsampled range of 7--12 M$_{\odot}$, stars no longer undergo AGB wind phases but end their lives instead as core collapse SNe. Where this transition occurs is uncertain, but is commonly taken to be 
%mm3
     at a mass $M_{\rm trans} \sim 
8$--10~M$_{\odot}$. This is a free parameter in our model; we take $M_{\rm trans} = 8$~M$_{\odot}$ as our fiducial value. In our model, stars below this mass eject their wind yields in an AGB phase only at the end of their lives, typically over a period comparable to or less than a few time steps ($\lesssim 10$ kyr). Stars above this mass are assumed to eject their stellar yields via line-driven stellar winds at a constant mass loss rate throughout their lifetime (neglecting Wolf-Rayet and luminous blue variable phases), ending their lives as a core-collapse SN (see Sect.~\ref{sec:stellar winds} for details on the wind energetics). Varying 
%mm3 $M_{\rm thresh}$ 
           $M_{\rm trans}$
changes both the time at which yields for stars around this mass are ejected (for reference, the lifetime of an 8 M$_{\odot}$ star is about 35--40~Myr), and the energy injection from these winds. \citet{Cote2017} explores how the choice of 
%mm3 $M_{\rm thresh}$
           $M_{\rm trans}$ affects galaxy abundances in a one-zone model. We neglect the effects of binary star evolution on stellar feedback, and discuss the significance of this in Sect.~\ref{sec:binary stars}.

@@

Type Ia SNe are an important additional source of galactic chemical enrichment. These iron group rich events are responsible for the $\sim$1 Gyr timescale turnover, or ``knee", in $[\alpha/\rm{Fe}]$ vs $[\rm{Fe}/\rm{H}]$ diagrams. We use the Type Ia SN yields given in \citet{Thielemann1986}, adopting a Type Ia SN model as discussed in Sect.~\ref{sec:Type Ia}. We emphasize that we only track Type Ia SNe occurring within the population of stars formed in this model, neglecting SNe from any possible pre-existing population, 
%mm3
     substantially limiting the number of Type Ia SNe occurring during the initial gigayear in 
     our models.
